---
title: "Quick introduction to R"
author: "Maria Doyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_notebook:
    toc: yes
    toc_depth: 3
    toc_float: yes
---


# Install R and RStudio
1. Install R from [here](https://cloud.r-project.org/)
2. Install RStudio from [here](https://rstudio.com/products/rstudio/download/#download)


# Install a package

To install a package in R (e.g. survival) on your computer, you can use `install.packages()`. You only need to install a package once per computer.
```{r}
install.packages('survival')
```

That is how to install packages from the CRAN repository. Many biological packages are in the Bioconductor repository, for those you install as below. You can Google the name of a package and install to see whether it's in CRAN or Bioconductor.

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("limma")
```


# Load a library

To use a package in an R session you use `library()`. You use library every time you start R and want to use a package. You only need to run it once in an R session.

```{r}
library(survival)
```

Library loads in all the functions in a package into your session. Sometimes we don't want to load in all functions, we just want to use one, in that case we can use  `name_of_the_package::name_of_the_function`.


# Import a file

To import a csv file you can use the built-in `read.csv()`. If it's a tab-delimited file you can use `read.delim()`. Or you if you have the tidyverse *readr* package installed you could use the `read_csv()` and `read_tsv()` functions, which are faster. To import an Excel file you can install the *readxl* package and use `read_excel()`.

```{r}
mydata <- read.csv("myfile.csv")
```


# To access a column 

You can use the `$` to access the contents of a column.

```{r}
mydata$column_name
```


# Data types

R uses data types, character (text), numeric (numbers), logical (TRUE/FALSE), factor (categories). Using factors is common in statistical analyses with R. To convert a column to data type factor you can do below.

```{r}
mydata$column_name <- factor(mydata$column_name) 
```


# Record versions used

It is good practice to note the versions of R and packages that you have used in your analysis with `sessionInfo()`.

```{r}
sessionInfo()
```


# Resources

[Introduction to R for biologists](https://melbournebioinformatics.github.io/r-intro-biologists/intro_r_biologists.html)  
[R for Data Science](https://r4ds.had.co.nz/)


